---
title: "Advanced Data Visualization with R"
author: "Abid Ali Shaikh"
date: "2025-02-05"
output: html_document
toc: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#setwd("~/../OneDrive/RSpace/2024-25/Advanced R/")
#install.packages("GGally")
#install.packages("gcookbook")
library(tidyverse)

```

## R Advanced Course

The Advanced visualization using ggplot in R course teaches several ways
we can draw scatterplots along-with the geom lines with varying
confidence intervals or without them using se=F. The libraries used
start with GGaly and gcookbook. You can learn more on
<http://abidalishaikh.github.io/quarto>.

Let's explore these two libraries; using data(package="gcookbook") shows
various datasets in this library as below: Data sets in package
'gcookbook':

-   aapl Apple stock data

-   anthoming Homing in desert ants

-   cabbage_exp Summary of cabbages data set

-   climate Global climate temperature anomaly data from 1800 to 2011

-   corneas Corneal thickness of eyes

-   countries Health and economic data about countries around the world
    from 1960-2010

-   drunk Convictions for drunkenness

-   heightweight Height and weight of schoolchildren

-   isabel Data from simulation of hurricane Isabel

-   madmen Successful sexual relations in Mad Men (TV show) madmen2
    Attempted sexual relations in Mad Men (TV show)\
    marathon Marathon and half-marathon times

-   pg_mean Means of results from an experiment on plant growth

-   plum Plum root cuttings (long format)

-   plum_wide Plum root cuttings (wide format)

-   simpledat Simple example data set

-   simpledat_long Simple example data set (long format)

-   tg Summarized ToothGrowth data

-   tophitters2001 Batting averages of the top hitters in Major League
    Baseball in 2001

-   uspopage Age distribution of population in the United States,
    1900-2002

-   uspopchange Change in population of states in the U.S. between 2000
    and 2010

-   wind Wind speed and direction worldpop World population estimates
    from 10,000 B.C. to 2,000 A.D.

Most of these datasets have longitudinal information over various time
periods.

*For this course, We read the already downloaded files from the previous
course of the specialization:*

```{r GGally}
library(tidyverse)
cces <- drop_na(read_csv("c:/Users/stech/OneDrive/RSpace/2024-25/RCourseggplot/cces_sample_coursera.csv"))
cel <- read_csv("c:/Users/stech/OneDrive/RSpace/2024-25/RCourseggplot/cel_volden_wiseman-_coursera.csv")

```

## Variables to consider

For our visualization we use educ and ideo5 of the cces data: both have
integer values the higher the educ level better the education and the
higher the ideo5 higher the strictness of ideology.

```{r educ vs ideo5 in cces data}
cor(cces$ideo5,cces$educ)
# we have -ve cor value, we can explore it using plot

cces |>
  select(ideo5,educ) |>
  ggplot(aes(x=ideo5, y=educ))+
  geom_jitter()

```

But we don't know its trend from the figure so we try add a geom_smooth
here, demonstrates the negative correlation trend.

```{r}
cces |>
  select(ideo5,educ) |>
  ggplot(aes(x=ideo5, y=educ))+
  geom_jitter()+
  geom_smooth() # we can add level=.9 for 90% C.I. or 
# we can also use se=F in geom_soomth to remove the C.I from the smooth line. also we can add method="lm" for straight abline.

```

`next we make a dataset with two correlated and one negatively correlated data and working on scatterplot matrix using GGally library`{=context}

```{r}
library(GGally)

var1 <- runif(100, min=0,max=100)
var2 <- var1 + rnorm(100,1,.2)
var3 <- var1*(-rnorm(100,1,.2))
df <- tibble(var1,var2,var3)
GGally::ggpairs(df)

```

ggpairs is robust; To control lower and diag layers' color and visual
aspects of ggpairs we make two functions

```{r upper lower ggpairs}

my_scatter <- function(data, mapping){
  ggplot(data,mapping=mapping)+
    geom_jitter(color="red")
}
my_density <- function(data,mapping){
  ggplot(data,mapping=mapping)+
    geom_density(alpha=.05, fill="darkblue")
}

ggpairs(df,
        lower=list(continuous = my_scatter),
        diag=list(continuous = my_density))


```

ggcorplot is another aspect of visualizing using shaded correlations

we work on pid7 (1 = Democ, 7: Repub) and pew_religion (1=important,
5=not imp.) variables

```{r ggcorplot}
library(ggcorrplot)
library(ggthemes)


 r<-  cces |>
  select(educ,pid7,pew_religimp)|>
  cor(use = "complete.obs")
  ggcorrplot(r,type = "lower",title = "ggcorplot",method = "circle",colors = c("brown","yellow","darkblue"),outline.color = "black")

  ggcorrplot(r,type="lower",ggtheme = theme_wsj())
  ggcorrplot(r,type="lower",ggtheme = theme_foundation())



  
```

## Working on higher legislative score

The next section visualizes the highest legislative (les) scores for
various congress members (thomas_name) in 114th congress. We consider
only 25 samples here.

```{r higher les score}
cel %>%
  filter (congress == 114) %>%
  slice_sample(n = 25)%>%
  select(les,thomas_name)%>%
  ggplot(aes(x=les, y=reorder(thomas_name,les)))+
  geom_point(size=3, color="steelblue")+
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(color="blue",linetype="dashed"))


```

## The geom_segment function

we flip the x and y in above plot without using coord_flip function,
here we flipped the x and y mappings alongside we also used the
geom_segment to plot a segment of the plot using x, xend and y, yend
parameters

```{r geom_segment}
cel %>%
  filter (congress == 114) %>%
  slice_sample(n = 25)%>%
  select(les,thomas_name)%>%
  ggplot(aes(y=les, x=reorder(thomas_name,les)))+
  geom_point(size=3, color="steelblue")+
  geom_segment(aes(x=thomas_name,xend=thomas_name,y=0,yend = les))+
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(color="blue",linetype="dashed"),
        axis.text.x = element_text(angle=90))



```

## Scatter Fitness

The following codes demonstrate some codes from the r-graphics.org using
the gcookbook library.

```{r gcookbook}
library(gcookbook)
heightweight %>%
  #select(ageYear,heightIn)%>%
  ggplot(aes(x=ageYear,y=heightIn,color=sex))+
  #geom_point()+
  geom_smooth(method="loess",level=.99)




```

## Maps

```{r maps}

library(maps)

wd <- map_data("world")
pakcities <- world.cities %>%
  filter((country.etc == "Pakistan" | country.etc == "India") & pop > 500000)

wd %>%
  filter(region == "Pakistan" | region=="India")%>%
  ggplot(aes(x=long,y=lat,group=group))+
  geom_polygon(fill="white",color="black")+
  geom_point(data=pakcities,aes(x=long,y=lat,group=NULL,size=pop),color="red")
  




```

## Animations using gganimate

```{r gganimate}
library(gganimate)
#library(transformr)
library(gifski)

mtcars  %>%
  ggplot(aes(x=factor(cyl),y=mpg))+
  geom_boxplot()+
  #facet_wrap(~gear, ncol=1)
  transition_states(gear)


cel <- read_csv("c:/Users/stech/OneDrive/RSpace/2024-25/RCourseggplot/cel_volden_wiseman-_coursera.csv")

cel %>%
  mutate(party = recode(dem, "1"="demo","0"="rep"))%>%
  group_by(year, party)%>%
  summarize("Seats" = n(),.groups = "drop")%>%
  ggplot(aes(x=year,y=Seats, fill=party))+
  geom_bar(stat="identity")+
  geom_hline(yintercept = 217)+
  transition_layers() # also use transition_time(years)

############ using layers for various geom filters
########### using all_pass seniority and party for 115 cong
cel115 <- filter(cel,cel$congress==115)
cel$party<-recode(cel$dem,`1`="Democrat",`0`="Republican")
d1 <- filter(cel,congress==115 & party=="Democrat")
d2 <- filter(cel,congress==115 & party=="Republican")
ggplot()+
  geom_jitter(aes(x=seniority,y=all_pass,color=party),data=d1)+
  geom_jitter(aes(x=seniority,y=all_pass,color=party),data=d2)+
  #geom_smooth(aes(x=seniority,y=all_pass,color=party),data=d1)+
  #geom_smooth(aes(x=seniority,y=all_pass,color=party),data=d2)+
  scale_color_manual(values=c("blue","red"))+
  transition_layers()

######################33333


ggplot(mtcars, aes(factor(cyl), mpg)) +
  geom_boxplot()

anim <- ggplot(mtcars, aes(factor(cyl), mpg)) +
  geom_boxplot() +
  transition_states(factor(cyl))

anim

anim2 <- anim + 
  enter_fade()+
  exit_fade()

anim2

cong_dat<- cel %>% 
  group_by(year,party) %>%
  summarise("Seats"=n())

anim3<-ggplot(cong_dat,aes(x=year,y=Seats,fill=party))+
  geom_bar(stat="identity")+
  geom_hline(yintercept=217)+
  scale_fill_manual(values=c("blue","red"))+
  transition_time(year)+
  shadow_wake(wake_length=1,alpha=FALSE,wrap=FALSE)

anim3
#anim_save("test.gif",animation=anim3)


```

## Plotly

```{r plotly}
library (plotly)

m<- mtcars  %>%
  ggplot(aes(x=factor(cyl),y=mpg))+
  geom_boxplot()+
  facet_wrap(~gear, ncol=1)

plotly::ggplotly(m)

dat <- data.frame(cond=rep(c("A","B"),each=10),
                  xvar = 1:20+rnorm(20,sd=3),
                  yvar = 1:20+rnorm(20,sd=3))
    
p<- dat %>%
  ggplot(aes(xvar,yvar))+
  geom_point(shape=1)

plotly::ggplotly(p)

######### Animation using ggplotly

cel <- read_csv("c:/Users/stech/OneDrive/RSpace/2024-25/RCourseggplot/cel_volden_wiseman-_coursera.csv")


p1 <-cel %>%
  mutate(Party = recode(dem,"1"="Democ","0"="Repub"))%>%
  ggplot(aes(y=les,x=seniority, frame=year))+
  geom_point(aes(color=Party))+
  scale_color_manual(values=c("red","blue"))

ggplotly(p1)

p1

```
